\chapter{VHbb Analysis Preamble}
\label{chap:vhbbpre}

\section{Overview}

The Higgs boson, discovered at the LHC in 2012, is predicted by the standard model to decay primarily to two $b$ quarks, with a branching factor of $0.582 \pm 0.007$ \cite{2016arXiv161007922D:2016:higgs}. 
Observation of this decay mode was recently reported by ATLAS \cite{HIGG-2018-04}. Whilst the dominant Higgs production mode at the LHC is gluon-gluon fusion, this mode has an overwhelming QCD multijet background and so sensitivity to the Higgs is low. The \hbb observation therefore searched for Higgs bosons produced in association with a vector boson (W or Z). This production mechanism results in leptonic final states from the decay of the vector boson, allowing for leptonic triggering, whilst at the same time significantly reducing the multi-jet background. 

A closely related analyses now searches for the \hbb decay of the Higgs boson, produced in association with a vector boson, when the vector boson and Higgs are highly boosted. The full Run-2 dataset is used for a total integrated luminosity of 139 fb$^{-1}$. The analysis is split into 0-, 1- and 2-lepton channels depending on the number of selected electrons and muons, to target the $\text{ZH} \rightarrow \nu\nu bb$, $\text{WH} \rightarrow \ell\nu bb$, $\text{ZH} \rightarrow \ell\ell bb$ processes, respectively, where $\ell$ is an electron or muon. In all channels, events are required to have exactly two $b$-tagged jets, which form the Higgs boson candidate. At least one of the $b$-tagged jets is required to have \pT greater than 45 GeV. Events are further split into 2-jet or 3-jet categories depending on whether additional, untagged jets are present. 

In the 0- and 1-lepton channels, the analysis is further split into signal and control regions. To leading order, there are no additional $b$-jets in the event other than the two coming from the reconstructed Higgs candidate. For this reason, there is a signal region veto (i.e. events are not accepted into the signal region) for events with additional $b$-tagged jets in the event. Events with additional $b$-tagged jets are included in the control region, which is highly pure in \ttbar events. The control region is used to constrain the normalisation of the \ttbar background.
